Empirical look at interstate migration PSID
========================================================

This is about PSID data 1994-2011.

```{r,echo=FALSE,message=FALSE}
library(data.table)
library(ggplot2)
# data prepared in ~/git/migration/migration-pkg/inst/makedata.r
load("~/git/migration/data/psid.RData")
library(survey)
des <- svydesign(id=~1,weights=~weight,data=dat)
subdes <- svydesign(id=~1,weights=~weight,data=dat[inter==TRUE])
subdesNomove <- svydesign(id=~1,weights=~weight,data=dat[inter==TRUE&why.moved!="not moved"])
subregdes <- svydesign(id=~1,weights=~weight,data=dat[interreg==TRUE])
subregdesNomove <- svydesign(id=~1,weights=~weight,data=dat[interreg==TRUE&why.moved!="not moved"])
# dat[inter==TRUE,image(table(state,state.l)>0,main="interstate moves. white is at least 1 moves")]
tmp <- dat[inter==TRUE,data.frame(expand.grid(to=sort(unique(State)),from=sort(unique(State.l))),movers=as.numeric(table(state,state.l)))]
tmp$movers <- cut(tmp$movers,breaks=c(-1,0,1,10,20),include.lowest=TRUE,labels=FALSE)
tmp$movers <- factor(tmp$movers,labels=c("0","1","1-10","11+"))
```

## Overview

Sample Characteristic | Value
--------------------- | -------------------
Number of unique inds | `r dat[,length(unique(pid))]`
Number of years       | `r dat[,length(unique(year))]`
range of years        | [`r dat[,range(year)]`]
Number of ind-years   | `r nrow(dat)`

## Statistics by year

All in current dollars.

```{r,echo=FALSE}
s <- svyby(~incomeFAM + own + Hvalue + mort1 + moveYES + inter + interreg,des,FUN=svymean,by=~year,na.rm=TRUE)
s <- s[,c(2,4,5,6,8,10,12)]
names(s) <- c("y","prop.own","HValue","MDebt","moved","mvd.state","mvd.region")
s
```

## Median Wealth and consumption by year

Current Dollars.

```{r,echo=FALSE,results='hide'}
dat[year<1999|year>2010,cons := 0]
dat[year<1999&year>1994,wealth := 0]
```

```{r,echo=FALSE}
print(dat[,list(assets=Hmisc::wtd.quantile(wealth,weights=weight,probs=0.5),cons=Hmisc::wtd.quantile(cons,weights=weight,probs=0.5)),by=year])
```

```{r echo=FALSE,results='hide'}
dat[year<1999|year>2010,cons :=NA]
dat[year<1999&year>1994,wealth := NA]
```



## how many periods do people stay in the sample?

```{r,results='hide',echo=FALSE}
dat[,diff(range(year)),by=pid][,barplot(table(V1),xlab="number of periods",ylab="number of individuals")]
```


## Moving patterns by state

```{r,echo=FALSE,fig.width=10,fig.height=10}
#Palette <- c(NA, colors()[c(81,655,556)])
Palette <- c(NA, "green","yellow","red")
ggplot(tmp,aes(x=from,y=to,z=movers,fill=movers)) + geom_tile() + scale_fill_manual(values=Palette) + theme(axis.text.x = element_text(angle=45,vjust=0.5) , panel.background = element_blank(), panel.grid.major = element_line(colour = "grey50"), panel.border= element_rect(color="black",fill=NA)) + ggtitle(sprintf("transition matrix of PSID interstate moves 1994-2011\nNumber of Moves: %d",dat[inter==TRUE,length(year)]))
```

## Moving patterns by Region

```{r,echo=FALSE}
#Palette <- c(NA, colors()[c(81,655,556)])
Palette <- c(NA, "green","blue","yellow","red","black")
tmp <- dat[interreg==TRUE,data.frame(expand.grid(to=sort(dat[,unique(region)]),from=sort(dat[,unique(region.from)])),movers=as.numeric(table(region,region.from)))]
tmp$movers <- cut(tmp$movers,breaks=c(-1,0,1,10,20,40,98),include.lowest=TRUE,labels=FALSE)
tmp$movers <- factor(tmp$movers,labels=c("0","1","1-10","11-20","21-40","41-97"))
ggplot(tmp,aes(x=from,y=to,z=movers,fill=movers)) + geom_tile() + scale_fill_manual(values=Palette) + theme(axis.text.x = element_text(angle=45,vjust=0.5) , panel.background = element_blank(), panel.grid.major = element_line(colour = "grey50"), panel.border= element_rect(color="black",fill=NA)) + ggtitle(sprintf("transition matrix of PSID interregional moves 1994-2011\nNumber of Moves: %d",dat[interreg==TRUE,length(year)]))
```

### How many interstate moves are interregional?

```{r,echo=FALSE}
prop.table(svytable(~inter+interreg,design=des),margin=1)
```


## Return Migration

* What is the empirical distribution of frequency of moves?

```{r,echo=FALSE,results='hide'}
dat[,list(freq=sum(interreg)),by=pid][freq>0,barplot(table(freq),ylab="number of households",xlab="number of interregional moves")]
````


* What is the average time between moves

```{r,echo=FALSE,results='hide'}
dat[interreg==TRUE,diff(year),by=pid][,barplot(table(V1),ylab="number of households",xlab="years between interregional moves")]
```

* what is the proportion of movers moving back to home region

```{r,echo=FALSE}
dat[interreg==TRUE,prop.table(table(region==Home.region))]
```

* What is the proportion of movers moving back to a previous state?

```{r,echo=FALSE}
ss[,prop.table(table(state.back))]
```

* proportion of movers back to previous region

```{r,echo=FALSE}
rr[,prop.table(table(region.back))]
```


## Determinants of interstate moves

### Sample Proportion Interstate Move by Age and Tenure

```{r,message=FALSE,echo=FALSE}
library(splines)
library(MASS)
#svymove <- svyby(~inter , by=~own + age,des,svymean)
#ggplot(subset(svymove,age>25&age<60),aes(x=age,y=interTRUE,color=own)) + geom_point()+ geom_smooth(method="lm",formula=y~ns(x,3))
ggplot(dat[age>25&age<60,list(interstate=mean(inter)),by=list(own,age)][order(age,own)],aes(x=age,y=interstate,color=own)) + geom_point() + geom_smooth(method="lm",formula=y~ns(x,3))
```

### Sample Proportion InterREGIONAL Move by Age and Tenure

```{r,message=FALSE,echo=FALSE}
#svymove <- svyby(~interreg , by=~own + age,des,svymean)
#ggplot(subset(svymove,age>25&age<60),aes(x=age,y=interregTRUE,color=own)) + geom_point()+ geom_smooth(method="lm",formula=y~ns(x,3))
ggplot(dat[age>25&age<60,list(interregional=mean(interreg)),by=list(own,age)][order(age,own)],aes(x=age,y=interregional,color=own)) + geom_point() + geom_smooth(method="lm",formula=y~ns(x,3))
```



### Self-reported reasons of ALL moves. % of population

```{r,message=FALSE,echo=FALSE}
options(scipen=3)
options(digits=4)
ftable(addmargins(svytable(~ why.moved + own,des,Ntotal=100)))
```

### Self-reported reasons of interstate moves. % of population who moved states

```{r,message=FALSE,echo=FALSE}
ftable(addmargins(svytable(~ why.moved + own,subdesNomove,Ntotal=100)))
```

### Self-reported reasons of interREGIONAL moves. % of population who moved regions

```{r,message=FALSE,echo=FALSE}
ftable(addmargins(svytable(~ why.moved + own,subregdesNomove,Ntotal=100)))
```

### Interstate Moves vs employment. % of population

```{r,message=FALSE,echo=FALSE}
ftable(addmargins(svytable(~ empstat + inter,des,Ntotal=100)))
```

### Interregional Moves vs employment. % of population

```{r,message=FALSE,echo=FALSE}
ftable(addmargins(svytable(~ empstat + interreg,des,Ntotal=100)))
```


### Interstate Moves vs marital status. % of population

```{r,message=FALSE,echo=FALSE}
ftable(addmargins(svytable(~ marstat + inter,des,Ntotal=100)))
```

### Interregional Moves vs marital status. % of population

```{r,message=FALSE,echo=FALSE}
ftable(addmargins(svytable(~ marstat + interreg,des,Ntotal=100)))
options(scipen=0)
options(digits=7)
```

## Probit: Determinants of all moves. 

```{r,warning=FALSE,echo=FALSE,message=FALSE,eval=FALSE}
library(erer)
library(texreg)
s=svyglm(formula=moveYES ~ age + I(age^2) + own + educ + incomeFAM + dnumkids + wealth + divorce,des,x=TRUE,family=quasibinomial(link="probit"))
Rsquared = 1-s$deviance/s$null.deviance
screenreg(list(s,maBina(s)),custom.model.names=c("estimates","marginal.effects"))
cat('Rsquared: ',Rsquared)
```

## Probit: Determinants of interstate moves. 

```{r,warning=FALSE,echo=FALSE,eval=FALSE}
s=svyglm(formula=inter ~ age + I(age^2) + own + educ + incomeFAM + dnumkids + wealth + divorce,des,x=TRUE,family=quasibinomial(link="probit"))
Rsquared = 1-s$deviance/s$null.deviance
screenreg(list(s,maBina(s)),custom.model.names=c("estimates","marginal.effects"))
cat('Rsquared: ',Rsquared)
```

## Probit: Determinants of interregional moves. 

```{r,warning=FALSE,echo=FALSE,eval=FALSE}
s=svyglm(formula=interreg ~ age + I(age^2) + own + educ + incomeFAM + dnumkids + wealth + divorce,des,x=TRUE,family=quasibinomial(link="probit"))
Rsquared = 1-s$deviance/s$null.deviance
screenreg(list(s,maBina(s)),custom.model.names=c("estimates","marginal.effects"))
cat('Rsquared: ',Rsquared)
```

## Average wage growth

Wage growth is higher for movers.

Sample |  growth movers (owner) | growth movers (renter)
------ | ------------------------- | -------------------------
all    | `r dat[incomeFAM>0&incomeFAM.l>0&own==TRUE,weighted.mean(log(incomeFAM)-log(incomeFAM.l),weight,na.rm=T)]` | `r dat[incomeFAM>0&incomeFAM.l>0&own==FALSE,weighted.mean(log(incomeFAM)-log(incomeFAM.l),weight,na.rm=T)]`
inter state    | `r dat[incomeFAM>0&incomeFAM.l>0&own==TRUE&inter==TRUE,weighted.mean(log(incomeFAM)-log(incomeFAM.l),weight,na.rm=T)]` | `r dat[incomeFAM>0&incomeFAM.l>0&own==FALSE&inter==TRUE,weighted.mean(log(incomeFAM)-log(incomeFAM.l),weight,na.rm=T)]`
all region   | `r dat[incomeFAM>0&incomeFAM.l>0&own==TRUE&interreg==TRUE,weighted.mean(log(incomeFAM)-log(incomeFAM.l),weight,na.rm=T)]` | `r dat[incomeFAM>0&incomeFAM.l>0&own==FALSE&interreg==TRUE,weighted.mean(log(incomeFAM)-log(incomeFAM.l),weight,na.rm=T)]`

## Linear Wage Equation

$$ \ln y_{it} = \beta_0 + \beta_1 \ln y_{it-1} + \beta_2 age_{it} + \beta_3 age^2_{it} + \beta_4 own_{it} + \beta_5 interreg_{it} + \gamma Z_{it} $$

```{r,echo=FALSE,warning=FALSE,results='hide',eval=FALSE}
dat[,dinc := log(incomeFAM)-log(incomeFAM.l)]
```
```{r,echo=FALSE,eval=FALSE}
des = svydesign(ids=~1,weights=~weight,data=dat[is.finite(dinc)&interreg==TRUE])
m = svyglm(dinc ~ age + I(age^2) + own + wealth + empstat+marstat + factor(numkids) + State + factor(educ),des)
screenreg(list(m),omit.coef="State|factor|empstat|marstat",digits=4)
cat('controlling for State, education, employment status, marital status and number of kids\nRsquare = ',1-m$deviance/m$null.deviance)
```


## Price movement over time

```{r,echo=FALSE,results='hide'}
vals = dat[,list(House.value=weighted.mean(Hvalue,weight,na.rm=T)),by=list(region,year)]
vals <- vals[complete.cases(vals)]
vals[,rank := rank(House.value),by=year]
```

```{r,echo=FALSE}
ggplot(vals,aes(x=year,y=House.value,color=region))+geom_line()
```

## Price ranking among states

```{r,echo=FALSE}
setkey(vals,year,region)
ggplot(vals,aes(x=year,y=region,z=factor(rank),fill=factor(rank))) + geom_tile()
```
